---
title: "Project-Farmers Market"
author: "Aakash Bolisetty"
date: "2/24/2020"
output: html_document
---
```{r libraries}
library(usmap)
library(ggplot2)
library(tidyr)
library(reshape2)
library(dplyr)
library(lubridate)
library(treemapify)
```

```{r setup data, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#creating a theme for plot
text_theme <- theme(
plot.title = element_text(color="black", size=14, face="bold.italic"),
axis.title.x = element_text(color="black", size=14, face="bold"),
axis.title.y = element_text(color="black", size=14, face="bold")
)
 
#creating a regions key
region_key <- data.frame(state.name, state.region,stringsAsFactors=FALSE)
region_key[nrow(region_key) + 1,] = c("District of Columbia","South")

#import the data set
fmarket <- read.csv("~/Downloads/fmarket.csv", na.strings="", stringsAsFactors=FALSE)

#dataframe with season 1 being split
fmarket<-fmarket%>%
  separate(Season1Date, c("StartS1", "EndS1"), " to ")

fmarket$DaysCount<-as.Date(fmarket$EndS1, format="%m/%d/%Y")-as.Date(fmarket$StartS1, format="%m/%d/%Y")

#based on years
fmarket_years <- fmarket%>%
  filter(!is.na(fmarket$DaysCount) & fmarket$DaysCount > 0)
fmarket_years$Year <- year(as.Date(fmarket_years$StartS1, format = "%m/%d/%Y"))

```

#Taking a count of the farmer markets across the United States of America from 2010-2020

```{r Farmer markets across the map}
farmers <- fmarket%>%
  group_by(State)%>%
  summarise(count=n())

farmers$fips <- fips(farmers$State)

plot_usmap(data = farmers, values = "count", color = "black",labels = T,label_color = 'white') + 
  scale_fill_continuous(name = "Farmer markets count", label = scales::comma) + 
  labs(title = "US States", subtitle = "Number of Farmer markets from 2010-2020")+
  theme(legend.position = "right")
#(plot)%>%ggsave("US Map Farmer Markets.png", . , dpi = 400, width = 8, height = 8)

```

#number of farmer markets per year

```{r farmer markets per year}
(fmarket_years %>%
  filter(Year!=2020)%>%
  group_by(Year)%>%
   summarise(Count=n())%>%
  select(Year,Count)%>%
mutate(Year_count=cumsum(Count))%>%
  ggplot(aes(x=factor(Year), y=Year_count, group=1)) +
  geom_line(color="red")+geom_point()+ labs(title = "Number of Farmer Markets", subtitle = "From 2011-2019") +xlab("Years") + ylab("Number of Farmer markets")+
  labs(color="Years")+text_theme)%>%ggsave("Farmer Market per Year.png", . , dpi = 400, width = 8, height = 4)
#(plot)%>%ggsave("Farmer Market per Year.png", . , dpi = 400, width = 8, height = 8)
```


#Statistics of the locations being used for the farmers markets

```{r Locations used}
fmarket%>%
  group_by(Location)%>%
  summarise(count=n())%>%
  drop_na()%>%
  ggplot(aes(area=count,label=Location,fill=Location))+geom_treemap()+ 
  geom_treemap_text(fontface = "italic", colour = "white", place = "centre",grow = FALSE)+
  ggtitle("Locations used for the markets")+text_theme
#(plot)%>%ggsave("Locations Used.png", . , dpi = 400, width = 8, height = 8)
```

#Visuals of the payment methods being used over the years based on region

```{r spread of payment methods}
payment_methods <- dplyr::select(fmarket,State,Credit,WIC,WICcash,SFMNP,SNAP)
payment_methods <- left_join(payment_methods,region_key,by = c("State" = "state.name"))
payment_methods <- select (payment_methods,-c(1))
payment_methods[-c(6)] <- lapply(payment_methods[-6], function(x) as.integer(x=="Y"))
payment_methods%>%
  group_by(state.region)%>%
  summarise(Credit=sum(Credit),WIC=sum(WIC),WICcash=sum(WICcash),SFMNP=sum(SFMNP),
            SNAP=sum(SNAP))%>%
   melt() %>% 
   filter(!is.na(state.region)) %>% 
   ggplot(aes(x=reorder(variable,-value),y=state.region, fill= value)) + 
   geom_tile()+
   scale_fill_gradient(low = "#56B1F7", high ="#132B43" )+
   geom_text(aes(label = value),color = "white") +
   ggtitle("Heatmap of Payment types") +  
   xlab("Payment types") + ylab("Regions")+
   labs(fill = "Number of users")+
   theme(panel.border = element_rect(colour = "darkblue", fill=NA, size=1))+text_theme
#(plot)%>%ggsave("Payment methods region.png", . , dpi = 400, width = 8, height = 8)
```

#payment methods through the years 2013 - 2019

```{r}
fmarket_years <- left_join(fmarket_years,region_key,by = c("State" = "state.name"))
fmarket_years <- fmarket_years[!is.na(fmarket_years$state.region),]
payment_methods_years <- dplyr::select(fmarket_years,state.region,Year,Credit,WIC,WICcash,SFMNP,SNAP)
payment_methods_years[-c(1:2)] <- lapply(payment_methods_years[-c(1:2)],
                                         function(x) as.integer(x=="Y"))
pm_years<-payment_methods_years%>%
  group_by(state.region,Year)%>%
  summarise(Credit=sum(Credit),WIC=sum(WIC),WICcash=sum(WICcash),SFMNP=sum(SFMNP),
            SNAP=sum(SNAP))%>%
  melt(id=c(1:2)) %>% 
  filter(Year>2012 & Year <2020)
pm_years$C_value <- ave(pm_years$value,paste0(pm_years$state.region,pm_years$variable),  FUN=cumsum)

ggplot(pm_years,aes(x=factor(Year),y=C_value,group=variable,color=variable))+
  geom_line()+ geom_point()+
  ggtitle("Payment types through the years") + 
  xlab("Years") + ylab("Number of Markets")+
  labs(color = "Payment methods")+
  theme(panel.border = element_rect(colour = "darkblue", fill=NA, size=1),
                     axis.text.x = element_text(angle=90))+
  facet_grid(~state.region,scales="free_y")+text_theme
#(plot)%>%ggsave("Payment methods through years.png", . , dpi = 400, width = 8, height = 8)
```

#Social media usage

```{r}
social_media <- dplyr::select(fmarket,Website,Facebook,Twitter,Youtube,OtherMedia)
social_media_lf <- gather(social_media,key="SocialMedia",value=measurement)
social_media_lf %>%
   filter(!is.na(measurement))%>%
   group_by(SocialMedia) %>%
  summarise(Count=n())%>%
  ggplot(aes(x=reorder(SocialMedia,-Count),y=Count,fill=SocialMedia))+
  geom_bar(stat="identity")+
  ggtitle("Social Media Usage") + 
  xlab("Social Media") + ylab("Number of Markets")+
  labs(fill = "Social Media")+
  theme(panel.border = element_rect(colour = "darkblue", fill=NA, size=1))+
  text_theme
  
#(plot)%>%ggsave("Payment methods through years.png", . , dpi = 400, width = 8, height = 4)
```



#Usage of social media over the years

```{r social media usage}
social_media_years <- dplyr::select(fmarket_years,Year,Website,Facebook,Twitter,Youtube,OtherMedia)
social_media_years[-c(1)] <- lapply(social_media_years[-c(1)],
                                         function(x) as.integer(!is.na(x)))
social_media_years <- social_media_years%>%
  group_by(Year) %>%
  filter(!Year<=2012 & !Year>2019)%>%
  drop_na() %>%
  summarise(Website=sum(Website),Facebook=sum(Facebook),
            Twitter=sum(Twitter),Youtube=sum(Youtube),OtherMedia=sum(OtherMedia))%>%
  melt(id="Year")
social_media_years$C_value <- ave(social_media_years$value,social_media_years$variable,  FUN=cumsum)

ggplot(social_media_years,aes(x=factor(Year),y=C_value,group=variable,color=variable))+
  geom_line()+geom_point()+
  ggtitle("Use of Social Media") + 
  xlab("Years") + ylab("Number of users")+labs(color="Social Media")+
  theme_dark()+
  theme(panel.border = element_rect(colour = "black", fill=NA, size=3))+text_theme
#(plot)%>%ggsave("Social media.png", . , dpi = 400, width = 8, height = 8)
```

#Sales of products in different regions

```{r}
product_sales <- left_join(fmarket,region_key,by = c("State" = "state.name"))
cols <- c(62, 30:59)
product_sales<-product_sales[,cols]
product_sales[-1]<-lapply(product_sales[-1],function(x) as.integer(x=="Y"))
product_sales%>%
  group_by(state.region)%>%
  summarise_if(is.numeric,sum)%>%
  drop_na()%>%
  melt(id="state.region")%>%
  ggplot(aes(fill=state.region, y=log(value), x=reorder(variable,-value),label = value)) + 
  geom_bar( stat="identity")+
  ggtitle("Region based sales for various products") + 
  xlab("Products") + ylab("Number of users(Log scale)")+labs(fill="Regions")+
  geom_text(color = "white",size = 3, position = position_stack(vjust = 0.5))+ 
  theme(panel.border = element_rect(colour = "darkblue", fill=NA, size=1),
        axis.text.x = element_text(angle=90))+coord_flip()+text_theme

#(plot)%>%ggsave("Product sales region.png", . , dpi = 400, width = 8, height = 8)
 
```

#number of days a farmer market remains open

```{r number of days for farmer markets}
fmarket_years %>%
  filter(fmarket_years$DaysCount <= 365)%>%
  ggplot(aes(x=DaysCount,fill=state.region))+geom_density(alpha=0.3)+
  ggtitle("Farmer market open days based on region") + 
  xlab("Number of days") + ylab("Density")+labs(fill="Regions")+
  facet_grid(.~state.region)+text_theme
#(plot)%>%ggsave("Farmer markets open days.png", . , dpi = 400, width = 8, height = 8)
```

